---
title: "Fetching LDC Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fetching LDC Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE   # network calls: not run during R CMD check
)
```

## Overview

`fetch_ldc_data()` retrieves LDC plot-level datasets from the
[Landscape Data Commons API](https://landscapedatacommons.org/) and
writes them as Tall CSV files that feed directly into
`generate_aero_inputs()`.

The function runs a **dependency chain**: it fetches `gap`, then filters
the returned PrimaryKeys to fetch `height`, then `lpi`, then `indicators`.
Only plots present across all datasets are kept in the final output.

## Prerequisites

```{r install, eval = FALSE}
# Install aeroinputs from GitHub
# install.packages("pak")
pak::pkg_install("Landscape-Data-Commons/aeroinputs")
```

You need LDC API credentials to access non-public data. Use 'trex::setup_keyring()' to store them securely in your system keyring. This only needs to be done once per user account and will allow you to access the API without hardcoding credentials in your scripts:

```r
username <- "your_email@example.com"
# Set up keyring, a prompt will ask you to enter your password securely
trex::setup_keyring(username)
# Store your password in the keyring (you will be prompted to enter it)
trex::store_password(username)
```

If you only want to access public data, you can skip this step and pass `username = NULL` to `fetch_ldc_data()`. However, note that some projects are restricted by default and will not be included in the results unless explicitly requested (see below).

## Basic usage

```{r basic}
library(aeroinputs)

username <- "your_email@example.com"
base_dir <- "~/LDC/AERO_test"

result <- fetch_ldc_data(
  username  = username,
  base_dir  = base_dir,
  n_rec     = 200,   # limit to 200 plots for a quick test
  write_out = TRUE,
  verbose   = TRUE
)
```

On success `result` is a list:

| Element       | Description                                      |
|---------------|--------------------------------------------------|
| `data`        | Named list: `gap`, `height`, `lpi`, `indicators`, `header` |
| `final_keys`  | Character vector of PrimaryKeys in all tables    |
| `out_dir`     | Path to the `Tall/` directory where CSVs were written |

## Using pre-selected PrimaryKeys

If you already know which plots you want, pass them via `primary_keys`.
The `excluded_project_keys` filter is skipped.

```{r keys}
my_keys <- c("key1", "key2", "key3")

result <- fetch_ldc_data(
  username     = username,
  primary_keys = my_keys,
  base_dir     = base_dir,
  write_out    = TRUE
)
```

## Filtering by ProjectKey

Use `project_keys` to fetch all plots belonging to one or more specific
projects. This overrides `excluded_project_keys` and is useful when you
want all plots from a known project without having to enumerate individual
`PrimaryKey`s.

```{r project-keys}
result <- fetch_ldc_data(
  username     = username,
  base_dir     = base_dir,
  project_keys = "Jornada_JERHM",
  write_out    = TRUE
)
```

You can pass multiple projects as a character vector:

```{r project-keys-multi}
result <- fetch_ldc_data(
  username     = username,
  base_dir     = base_dir,
  project_keys = c("Jornada_JERHM", "BLM_AIM"),
  write_out    = TRUE
)
```

## Restricted projects and default exclusions

Some projects are **restricted by default** because their data is pending
publication or currently under review and should not be used in general
analyses:

- **`NWERN_*`** — all National Wind Erosion Research Network sites
  (e.g. `NWERN_Moab`, `NWERN_JER`, `NWERN_CPER`, etc.) are excluded
  pending formal data publication.
- **`CFO_USGS`** — excluded because the dataset is currently under
  review and has not yet been cleared for unrestricted use.

These exclusions are applied **automatically** in all calls unless you
explicitly override them via `project_keys` or `primary_keys` (see
priority table below).

```{r restricted}
# This will work: project_keys bypasses the default exclusion
result <- fetch_ldc_data(
  username     = username,
  base_dir     = base_dir,
  project_keys = "CFO_USGS",   # explicitly requested: exclusion is bypassed
  write_out    = TRUE
)
```

### Priority and exclusion logic

The three filter arguments follow a strict priority order:

| Argument                | Priority | Behaviour |
|-------------------------|----------|-----------|
| `primary_keys`          | Highest  | Uses exactly those keys; ignores `project_keys` and `excluded_project_keys` (including mandatory exclusions) |
| `project_keys`          | Middle   | Fetches all PrimaryKeys for those projects; ignores `excluded_project_keys` (including mandatory exclusions) |
| `excluded_project_keys` | Lowest   | Applied only when neither `primary_keys` nor `project_keys` is supplied. **`NWERN_*` and `CFO_USGS` are always appended**, even when you supply your own list |

Examples of how exclusions interact:

| Call | Excluded projects |
|------|------------------|
| `fetch_ldc_data(...)` | All `NWERN_*` + `CFO_USGS` (default) |
| `fetch_ldc_data(..., excluded_project_keys = "BLM_AIM")` | `BLM_AIM` + all `NWERN_*` + `CFO_USGS` |
| `fetch_ldc_data(..., excluded_project_keys = character(0))` | All `NWERN_*` + `CFO_USGS` (mandatory, cannot be removed this way) |
| `fetch_ldc_data(..., project_keys = "CFO_USGS")` | None — `project_keys` bypasses all exclusions |
| `fetch_ldc_data(..., primary_keys = my_keys)` | None — `primary_keys` bypasses all exclusions |

> **Note:** The only way to access `NWERN_*` or `CFO_USGS` data is to
> request them explicitly via `project_keys` or `primary_keys`.
> Supplying them in `excluded_project_keys = character(0)` will **not**
> remove the mandatory exclusion.

## Caching the header

The full LDC header table (~65 k rows) is cached as an `.RData` file
at `header_cache_file`. On subsequent calls the cache is loaded instead
of hitting the API again, saving ~30 seconds per run.

```{r cache}
# Force a fresh header download by deleting the cache:
# file.remove(file.path(base_dir, "header.RData"))

result <- fetch_ldc_data(
  username          = username,
  base_dir          = base_dir,
  header_cache_file = file.path(base_dir, "header.RData"),
  write_out         = TRUE
)
```

## Logging non-public keys

When the API returns non-public data warnings, the offending PrimaryKeys
are appended to `nonpublic_file` (a CSV). Set `nonpublic_file = NULL` to
silence logging.

```{r log}
result <- fetch_ldc_data(
  username       = username,
  base_dir       = base_dir,
  nonpublic_file = file.path(base_dir, "nonpublic.csv"),
  write_out      = TRUE
)
```

## Output files

When `write_out = TRUE`, the following files are created under `out_dir`
(`<base_dir>/Tall/` by default):

```
Tall/
  header.csv
  gap_tall.csv
  height_tall.csv
  lpi_tall.csv
  geoIndicators.csv
```

These files are the required inputs for `generate_aero_inputs()`.
