---
title: "Fetching LDC Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fetching LDC Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE   # network calls: not run during R CMD check
)
```

## Overview

`fetch_ldc_data()` retrieves LDC plot-level datasets from the
[Landscape Data Commons API](https://landscapedatacommons.org/) and
writes them as Tall CSV files that feed directly into
`generate_aero_inputs()`.

The function runs a **dependency chain**: it fetches `gap`, then filters
the returned PrimaryKeys to fetch `height`, then `lpi`, then `indicators`.
Only plots present across all datasets are kept in the final output.

## Prerequisites

```{r install, eval = FALSE}
# Install aeroinputs from GitHub
# install.packages("pak")
pak::pkg_install("Landscape-Data-Commons/aeroinputs")
```

You need LDC API credentials. Store them as environment variables to
avoid hard-coding credentials in scripts:

```r
# In your .Renviron (see usethis::edit_r_environ()):
# API_USERNAME=your_username
# API_PASSWORD=your_password
```

## Basic usage

```{r basic}
library(aeroinputs)

# Obtain an API token
my_token <- trex::get_ldc_token(
  username = Sys.getenv("API_USERNAME"),
  password = Sys.getenv("API_PASSWORD")
)

base_dir <- "~/LDC/AERO_test"

result <- fetch_ldc_data(
  token     = my_token,
  base_dir  = base_dir,
  n_rec     = 500,   # limit to 500 plots for a quick test
  n_offset  = 1,
  write_out = TRUE,
  verbose   = TRUE
)
```

On success `result` is a list:

| Element       | Description                                      |
|---------------|--------------------------------------------------|
| `data`        | Named list: `gap`, `height`, `lpi`, `indicators`, `header` |
| `final_keys`  | Character vector of PrimaryKeys in all tables    |
| `out_dir`     | Path to the `Tall/` directory where CSVs were written |

## Using pre-selected PrimaryKeys

If you already know which plots you want, pass them via `primary_keys`.
The `excluded_project_keys` filter is skipped.

```{r keys}
my_keys <- c("key1", "key2", "key3")

result <- fetch_ldc_data(
  token        = my_token,
  primary_keys = my_keys,
  base_dir     = base_dir,
  write_out    = TRUE
)
```

## Caching the header

The full LDC header table (~65 k rows) is cached as an `.RData` file
at `header_cache_file`. On subsequent calls the cache is loaded instead
of hitting the API again, saving ~30 seconds per run.

```{r cache}
# Force a fresh header download by deleting the cache:
# file.remove(file.path(base_dir, "header.RData"))

result <- fetch_ldc_data(
  token             = my_token,
  base_dir          = base_dir,
  header_cache_file = file.path(base_dir, "header.RData"),
  write_out         = TRUE
)
```

## Logging non-public keys

When the API returns non-public data warnings, the offending PrimaryKeys
are appended to `nonpublic_file` (a CSV). Set `nonpublic_file = NULL` to
silence logging.

```{r log}
result <- fetch_ldc_data(
  token          = my_token,
  base_dir       = base_dir,
  nonpublic_file = file.path(base_dir, "nonpublic.csv"),
  write_out      = TRUE
)
```

## Output files

When `write_out = TRUE`, the following files are created under `out_dir`
(`<base_dir>/Tall/` by default):

```
Tall/
  header.csv
  gap_tall.csv
  height_tall.csv
  lpi_tall.csv
  geoIndicators.csv
```

These files are the required inputs for `generate_aero_inputs()`.
