---
title: "Fetching SOLUS Soil Texture Rasters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fetching SOLUS Soil Texture Rasters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE   # network calls: not run during R CMD check
)
```

## Overview

`fetch_solus()` streams SOLUS 100 soil-property rasters from Google
Cloud Storage. You can retrieve a full CONUS raster grid, crop it to a
spatial area of interest, or convert the extracted values to an `aqp`
`SoilProfileCollection`.

For the AERO workflow you typically need only the `sandtotal` and
`claytotal` layers at the `0` cm depth slice.

## Basic grid retrieval

```{r grid}
library(aeroinputs)

base_dir <- "~/LDC/AERO_test"
outfile  <- file.path(base_dir, "soil_texture_w_sand_frac.tif")

# Download the full SOLUS index and build a streaming raster
solus_grid <- fetch_solus(return = "grid")

# Subset to the two layers needed for AERO
soil_texture <- solus_grid[[c("sandtotal_0_cm_p", "claytotal_0_cm_p")]]

# Write to disk (this streams from remote storage - may take a few minutes)
terra::writeRaster(soil_texture, filename = outfile, overwrite = TRUE)
```

> **Note:** The full SOLUS raster covers CONUS and can be several GB.
> Write it to disk once and reuse; do not re-download on every session.

## Requesting specific variables and depths

```{r subset}
# All available sandtotal depth slices
sol <- fetch_solus(
  variables    = "sandtotal",
  depth_slices = c(0, 5, 15, 30, 60, 100, 150),
  output_type  = "prediction",
  return       = "grid"
)
terra::names(sol)
```

## Cropping to an area of interest

Pass any `sf`, `SpatVector`, or `SpatRaster` as `x` to crop the result
before downloading.

```{r crop}
library(sf)

# Example: crop to a bounding box in the Colorado Plateau
aoi <- sf::st_bbox(c(xmin = -112, xmax = -109, ymin = 36, ymax = 38),
                   crs = 4326) |>
  sf::st_as_sfc() |>
  sf::st_as_sf()

sol_crop <- fetch_solus(
  x            = aoi,
  variables    = c("sandtotal", "claytotal"),
  depth_slices = 0,
  output_type  = "prediction",
  return       = "grid"
)
```

## Extracting a SoilProfileCollection

Set `return = "spc"` to extract raster values at sample points and
convert them to an `aqp::SoilProfileCollection`.

```{r spc}
sol_spc <- fetch_solus(
  x            = aoi,
  variables    = c("sandtotal", "claytotal", "ph1to1h2o"),
  depth_slices = c(0, 5, 15, 30, 60, 100),
  output_type  = "prediction",
  return       = "spc",
  samples      = 100,
  method       = "linear"
)

aqp::plot(sol_spc, color = "sandtotal_p")
```

## Available variables

The 20 SOLUS 100 properties currently available:

```
anylithicdpt  caco3    cec7      claytotal  dbovendry
ec            ecec     fragvol   gypsum     ph1to1h2o
resdept       sandco   sandfine  sandmed    sandtotal
sandvc        sandvf   sar       silttotal  soc
```

## Interpolation methods

When `return = "spc"`, values are interpolated between depth slices.
Available methods (from `stats::approxfun` / `stats::splinefun`):

| Method      | Description                       |
|-------------|-----------------------------------|
| `"slice"`   | Each depth slice as a 1 cm slice  |
| `"step"`    | SOLUS native depth intervals      |
| `"linear"`  | Linear interpolation (recommended)|
| `"constant"`| Step-function interpolation       |
| `"fmm"`     | Fritsch-Carlson spline            |
| `"natural"` | Natural cubic spline              |

Use `method = "slice"` or `"step"` when you want to preserve the raw
SOLUS predictions without modification.
