---
title: "Generating AERO Inputs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating AERO Inputs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE   # requires local data files
)
```

## Overview

`generate_aero_inputs()` converts the LDC Tall CSVs produced by
`fetch_ldc_data()` and a soil-texture GeoTIFF from `fetch_solus()` into
the per-plot files that the AERO wind-erosion model requires:

| Output                  | Description                                     |
|-------------------------|-------------------------------------------------|
| `gap/<PK>.txt`          | Canopy gap distances (fraction) for each plot   |
| `<PK>.ini`              | AERO INI configuration for each plot            |
| `input_data.csv`        | Combined summary table of all plot inputs       |

## Prerequisites

You need the outputs of both `fetch_ldc_data()` and `fetch_solus()`.
See the corresponding vignettes for how to produce them.

```
<base_dir>/
  Tall/
    header.csv
    gap_tall.csv
    height_tall.csv
    lpi_tall.csv
  soil_texture_w_sand_frac.tif   # from fetch_solus()
```

## Basic usage

```{r basic}
library(aeroinputs)

base_dir     <- "~/LDC/AERO_test"
output_dir   <- file.path(base_dir, "aero_inputdata")
texture_file <- file.path(base_dir, "soil_texture_w_sand_frac.tif")

result <- generate_aero_inputs(
  data_dir     = base_dir,
  output_dir   = output_dir,
  texture_file = texture_file,
  write_out    = TRUE,
  verbose      = TRUE
)
```

The function returns a list:

| Element        | Description                                         |
|----------------|-----------------------------------------------------|
| `input_data`   | Data frame combining texture, bare soil, height, gap |
| `plots_texture`| Data frame with extracted sand/clay fractions        |
| `files_written`| Character vector of all paths written to disk        |

## Texture raster requirements

The raster at `texture_file` must have exactly two layers named (or
containing) `"sand"` and `"clay"`, in **percent** (0-100). The function
divides by 100 internally before writing fractions to the INI files.

```{r texture-check}
library(terra)

r <- rast(texture_file)
names(r)   # should include "sandtotal_0_cm_p" and "claytotal_0_cm_p"
nlyr(r)    # 2 after subsetting in fetch_solus example
```

## Inspecting outputs

```{r inspect}
# Summary table
head(result$input_data)

# List gap files written
head(result$files_written)

# Read one INI file
readLines(file.path(output_dir, result$plots_texture$PK_texture[1]))
```

## Using write_out = FALSE

Set `write_out = FALSE` to run the pipeline in memory without touching
disk, for example during development or in unit tests.

```{r no-write}
result <- generate_aero_inputs(
  data_dir     = base_dir,
  texture_file = texture_file,
  write_out    = FALSE
)

# input_data is NULL when write_out = FALSE
is.null(result$input_data)  # TRUE
nrow(result$plots_texture)  # plots processed
```

## Complete workflow

```{r workflow}
# 1. Fetch soil texture raster (once)
fetch_solus(
  variables    = c("sandtotal", "claytotal"),
  depth_slices = 0,
  output_type  = "prediction",
  return       = "grid",
  filename     = texture_file,
  overwrite    = FALSE
)

# 2. Fetch LDC data (once)
fetch_ldc_data(
  token    = my_token,
  base_dir = base_dir,
  n_rec    = 500
)

# 3. Build AERO inputs
result <- generate_aero_inputs(
  data_dir     = base_dir,
  output_dir   = output_dir,
  texture_file = texture_file
)
```
